Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-07-31 paper.tex
Update paper.tex

This paper investigates the microscopic parameters that control HDFG output. \cite{Bonn2024, McDonnell2024}
In \autoref{steadystate}, we first identify the vibrational selection rules. 
The selection rules of electronically-resonant HDFG are then identified through a Herzberg-Teller dipole expansion.
HDFG is found to be allowed in any harmonic system that has infrared active vibrations.
A simple harmonic oscillator model system is used to simulate HDFG spectra and illustrate how the excited state potential energy surface affects the hyper-Raman excitation spectrum.
After developing selection rules, some quantitative aspects of HDFG are discussed in \autoref{quant}.
The main findings are summarized in \autoref{conclusion}.

@@

Can vSFG experimental setups detect HDFG signals?
SFG is often consider weak compared to third order spectroscopies because it is usually surface selective,\cite{RN132, RN133} so vSFG detection setups are often quite sensitive.
However, HDFG is a weak $\chi^{(3)}$ process because it depends upon hyper-Raman cross-sections, which are commonly several orders of magnitude weaker than corresponding Raman transitions.\cite{RN515}
To evaluate the feasibility of HDFG for vSFG practitioners, we have compared the relative output of each process using nominal parameters. 
As calculated in \autoref{appendixA}, HDFG's weak $\chi^{(3)}$ value and SFG's low emission cancel somewhat, so that the output polarization of HDFG and vSFG are comparable ($\abs{P_\text{vSFG} / P_\text{HDFG}} \sim 1$).
Transient HDFG, or IR-pump-HDFG-probe, analogous to IR-pump-vSFG-probe, is therefore also feasible under similar experimental conditions, and provides vSFG practitioners a simple method to measure bulk dynamics without significant changes in an experimental setup. 
Therefore, HDFG is a feasible spectroscopy for practitioners of vSFG both in terms of experimental setup and detection feasibility.

@@

Using \autoref{eq:nfgamma} and \autoref{sivebeta}, we can write
\begin{equation}\label{chi3}
		\chi^{(3)}_{IJKL} = -\frac{NF}{4D \hbar \varepsilon_0 \Delta_{gv}} \langle \beta_{ijk} \mu_l \rangle \rho_{gg}\\
\end{equation}
For simplicity, we take $\rho_{gg} = 1$.
The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}
Briefly, a tensor in the molecular frame, $A_{ijkl}$, is transformed into an element of the same tensor in the laboratory frame, $\langle A \rangle_{IJKL}$, through $\langle A \rangle_{IJKL}$ = $\theta^{ijkl}_{IJKL} A_{ijkl}$, where summation over repeated indices is implied and $\theta$ is the transformation operator. \cite{McDonnell2024}
Orientational averaging shows specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. \cite{Bersohn1966, Willetts1992, Kauranen1996}
Substituting in \autoref{SIVEselection}, we see
% DDK: why does it matter that chi is imaginary?
\begin{equation}\label{betasive}
	\left\langle \frac{\partial \beta}{\partial Q_v} {\frac{\partial \mu}{\partial Q_v}} \right\rangle _{IJKL} = -\frac{8D \varepsilon_0}{NF}  {\Gamma_{gv} \omega_{vg}} \chi^{(3)}_{IJKL}
\end{equation}
$\chi^{3}$ itself can be extracted with use of an internal standard.\cite{Levenson1974_1, Levenson1974_2}
Since $\abs{\partial \mu / \partial Q}$ values can be extracted from FT-IR spectra,\cite{RN459} HDFG can give quantitative information on the magnitude of $\beta_{ijk}$ for infrared active vibrations.
Additionally, by using non-degenerate input frequencies to stimulate the hyper-Raman transition, the asymmetric properties of $\beta_{ijk}$ can be examined and quantified. \cite{Christie1971, Denisov1986, Kozich2007}
These quantitative aspects of HDFG could prove useful for examining how well computational methods calculate $\beta_{ijk}$ values.

>>>>

This paper investigates the microscopic parameters that control HDFG output. \cite{Bonn2024, McDonnell2024}
In \autoref{steadystate}, we first identify the vibrational selection rules. 
The selection rules of electronically-resonant HDFG are then identified through a Herzberg-Teller dipole expansion.
HDFG is found to be allowed in any harmonic system that has infrared active vibrations.
A simple harmonic oscillator model system is used to simulate HDFG spectra and illustrate how the excited state potential energy surface affects the hyper-Raman excitation spectrum.
After developing selection rules, some quantitative aspects of HDFG are discussed in \autoref{quant}.

@@

Can vSFG experimental setups detect HDFG signals?
SFG setups typically leverage the surface selectivity of $\chi^{(2)}$ polarizations,\cite{RN132, RN133} which results in extremely small emission volumes, so SFG setups are typically quite sensitive.
However, HDFG is a weak $\chi^{(3)}$ process because it depends upon hyper-Raman cross-sections, which are commonly several orders of magnitude weaker than corresponding Raman transitions.\cite{RN515}
To evaluate the feasibility of HDFG for vSFG practitioners, we have compared the relative output of each process using nominal parameters. 
As calculated in \autoref{appendixA}, HDFG's weak $\chi^{(3)}$ value and SFG's low emission cancel somewhat, so that the output polarization of HDFG and vSFG are comparable ($\abs{P_\text{vSFG} / P_\text{HDFG}} \sim 1$).
Transient HDFG, or IR-pump-HDFG-probe, analogous to IR-pump-vSFG-probe, is therefore also feasible under similar experimental conditions, and provides vSFG practitioners a simple method to measure bulk dynamics without significant changes in an experimental setup. 
Therefore, HDFG is a feasible spectroscopy for practitioners of vSFG both in terms of experimental setup and detection feasibility.

@@

Using \autoref{eq:nfgamma} and \autoref{sivebeta}, we can write
\begin{equation}\label{chi3}
		\chi^{(3)}_{IJKL} = -\frac{NF}{4D \hbar \varepsilon_0 \Delta_{gv}} \langle \beta_{ijk} \mu_l \rangle \rho_{gg}\\
\end{equation}
For simplicity, we take $\rho_{gg} = 1$.
The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}
Briefly, a tensor in the molecular frame, $A_{ijkl}$, is transformed into an element of the same tensor in the laboratory frame, $\langle A \rangle_{IJKL}$, through $\langle A \rangle_{IJKL}$ = $\theta^{ijkl}_{IJKL} A_{ijkl}$, where summation over repeated indices is implied and $\theta$ is the transformation operator. \cite{McDonnell2024}
Orientational averaging shows specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. \cite{Bersohn1966, Willetts1992, Kauranen1996}
Substituting in \autoref{SIVEselection}, we see
% DDK: why does it matter that chi is imaginary?
\begin{equation}\label{betasive}
	\left\langle \frac{\partial \beta}{\partial Q_v} {\frac{\partial \mu}{\partial Q_v}} \right\rangle _{IJKL} = -\frac{8D \varepsilon_0}{NF}  {\Gamma_{gv} \omega_{vg}} \chi^{(3)}_{IJKL}
\end{equation}
$\chi^{(3)}$ itself can be extracted with use of an internal standard.\cite{Levenson1974_1, Levenson1974_2}
Since $\abs{\partial \mu / \partial Q}$ values can be extracted from FT-IR spectra,\cite{RN459} HDFG can give quantitative information on the magnitude of $\beta_{ijk}$ for infrared active vibrations.
Additionally, by using non-degenerate input frequencies to stimulate the hyper-Raman transition, the asymmetric properties of $\beta_{ijk}$ can be examined and quantified. \cite{Christie1971, Denisov1986, Kozich2007}
These quantitative aspects of HDFG could prove useful for examining how well computational methods calculate $\beta_{ijk}$ values.
